\ttitle

### 2013-05-28 Appendices/AppendixA-summaries.tex
Summary Direct Density Ratio Estimation

%----------------------------------------------------------------------------------------
\section{Support Vector Machines}

@@

%----------------------------------------------------------------------------------------
\section{CUSUM for variance}

@@

>>>>

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Support Vector Machines}

@@

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{CUSUM for variance}

@@

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Density Ratio Estimation}

\subsection{Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation}
Kawahara and Sugiyama \cite{kawahara2009change}, 2009, 45 refs.

``This paper provides a change-point detection algorithm based on direct density-ratio estimation that can be computed very efficiently in an online manner''.

%--------------------------------------------
\subsubsection{Introduction}
The problem of change-point detection is well studied over the last decades in the field of statistics.
A common statistical formulation of change-point detection is to consider the probability distributions over past and present data intervals, and regard the target time as a change-point if the two distributions are significantly different.
Some approaches (such as CUSUM and GLR) make use of the \emph{log-likelihood ratio}, and are extensively explored in the data mining community.
Many approaches (novelty detection, maximum-likelihood ratio, learning of autoregressive models, subspace identification) rely on pre-specified parametric models (probability density models, autoregressive models, state-space models).
That makes it less applicable to real-life problems.
There have been some non-parametric density estimation approaches proposed, but that is known to be a hard problem.
The key idea of this paper to directly estimate the \emph{ratio} of the probability densities (also known as \emph{importance}).
The Kullback-Leibler Importance Estimation Procedure (KLIEP) is an example, but it is a batch algorithm.
This paper introduces an online version of the KLIEP algorithm and develops a flexible and computationally efficient change-point detection method.
This method is equipped with a natural cross validation procedure and thus the value of tuning parameters can be objectively determined.

\subsubsection{Problem formulation and Basic Approach}
Let $\vectorsym{y}(t) (\in \mathbb{R}^d)$ be a $d$-dimensional time series sample at time $t$.
The task is to detect whether there exists a change point between two consecutive time intervals, called the \emph{reference} and \emph{test} intervals.
The conventional algorithms consider the likelihood ratio over samples from the two intervals.
Since time-series samples generally are not independent over time it is hard to deal with them directly.
To overcome this difficulty, we consider \emph{sequences} of samples in the intervals: $\matrixsym{Y}(t)(\in \mathbb{R}^{dk})$ is the forward subsequence of length $k$ at time $t$.
This is a common practice in subspace identification since it takes implicitly time correlation into consideration.
The algorithm in the paper is based on the logarithm of the likelihood ration of the \emph{sequence sample} $\matrixsym{Y}$:
\begin{equation}
  s(\matrixsym{Y}) = \text{ln}\frac{p_\text{te}(\matrixsym{Y})}{p_\text{rf}(\matrixsym{Y})}
\end{equation}
where $p_{\text{te}}(\matrixsym{Y})$ and $p_{\text{rf}}(\matrixsym{Y})$ are the probability density functions of the reference and test sequence samples.
Let $t_{\text{rf}}$ and $t_{\text{te}}$ be the starting points of the reference and test intervals.
Decide if there is a change-point between the reference and test interval by monitoring the logarithm of the likelihood ratio:
\begin{equation}
  S = \sum_{i=1}^{n_\text{te}} \text{ln} \frac{p_{\text{te}}(\matrixsym{Y}_\text{te}(i))}{p_\text{rf}(\matrixsym{Y}_\text{te}(i))}
\end{equation}
If, for a predefined $\mu > 0$, it holds that $S > \mu$ then a change occurs.
The remaining question is how to calculate the density ratio, because it is unknown and we need to estimate it from examples.
The naive approach would be to first estimate the densities for the reference and test interval separately and then take the ratio.
This approach via non-parametric density estimation may not be effective --- directly estimating the density ratio without estimating the densities would be more promising.

The direct estimation of the density ratio is based on the Kullback-Leibler Importance Estimation Procedure.
Let us model the density ratio $w(\matrixsym{Y})$ by a non-parametric Gaussian kernel model:
\begin{equation}
  \hat{w}(\matrixsym{Y}) = \sum_{l=1}^{n_\text{te}} \alpha_l K_\sigma(\matrixsym{Y}, \matrixsym{Y}_\text{te}(l)),
\end{equation}
where $\{\alpha_l\}_{l=1}^{n_\text{te}}$ are parameters to be learned from the data samples and $K_\sigma(\matrixsym{Y}, \matrixsym{Y}')$ is the Gaussian kernel function with mean $\matrixsym{Y}'$ and standard deviation $\sigma$.

\subsubsection{Online Algorithm}
...
